---
title: "ESM 204, Homework #2"
author: "Drew Andersen"
date: "4/25/2022"
output: 
  html_document:
    code_folding: hide
---
```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(here)
library(tidyverse)
library(kableExtra)
```
## Introduction
- For this exercise, I will simulate a water market for four irrigation districts.
- Current/baseline water usage is given for each district.
- Prices are in $/Acre Foot and quantities are in Acre Feet (AF).

## Question 1
```{r}
#Loading in the data and baselines

w_data <- read_csv(here("Water_Districts.csv"))[,-1] %>%  #upload and drop first column
  rename(Kern = MC_Kern, Ventura = MC_Ventura, Mojave = MC_Mojave, Antelope = MC_Antelope)

mc_water <- pivot_longer(w_data, !Reduction, names_to = "District", values_to = "MAC") #pivot table to create column for districts (populated by four districts)

w_base_kern = 150 # Baseline water usage in Acre Ft. These values are E0 for each district
w_base_moj = 140
w_base_ant = 220
w_base_vent = 245 

agg_base = w_base_kern + w_base_moj + w_base_ant + w_base_vent
# Quantity is Acre Feet and Price is $/Acre Foot

w_abated = 255 #Given by a reduction from 755 AF --> 500 AF
```

```{r}
#Linear Regressions of data sets to get slope of each abatement curve

kern_slope <- lm(Kern ~ 0 + Reduction, data = w_data)$coef[1]
moj_slope <- lm(Mojave ~ 0 + Reduction, data = w_data)$coef[1]
ant_slope <- lm(Antelope ~ 0 + Reduction, data = w_data)$coef[1]
vent_slope <- lm(Ventura ~ 0 + Reduction, data = w_data)$coef[1]

agg_a_slope = 1/((1/kern_slope) + (1/moj_slope) + (1/ant_slope) + (1/vent_slope))
```

```{r, fig.align="center"}
#plotting abatement curves
ggplot(data = mc_water, aes (x = Reduction, y = MAC, color = District)) +
  geom_line() +
  geom_abline(intercept = 0, slope = kern_slope, color = "blue") +
  geom_abline(intercept = 0, slope = vent_slope, color = "orange") +
  geom_abline(intercept = 0, slope = moj_slope, color = "green") + 
  geom_abline(intercept = 0, slope = ant_slope, color = "red") +
#geom_abline(intercept = 0, slope = agg_a_slope, color = "purple") +

  scale_color_manual(values = c("red","blue","green","orange")) +
  xlim(0,300) +
  ylim(0,400) +
  theme_minimal() +
  labs(title = "Marginal Abatement Cost Curves by Water District", y = "Marginal Abatement Cost ($/AF)", x = "Quantity Abated (AF)") +
  theme(legend.position = c(.85,.75))
```

## Question 2
Using the baseline water use of each district (E0), the marginal abatement curves (following the form P = M * Q, where M is each districts positive slope and Q is the quantity abated), and the equation A = E0 - E (where A is the quantity abated and E is quantity used post-abatement), one can derive the demand curves for each district using basic algebra:

P = M * Q

P = M * (E0 - E)

P = (M * E0) - (M * E)

```{r}
#Deriving Demand Functions & their inverses - inverse functions will be used in part 3c

kern_demand <- function(emiss) { 
  demand <- (kern_slope * w_base_kern) - (kern_slope * emiss)
  return(demand)
}
kern_demand_inv <- function(demand) {
  emiss <- w_base_kern - (demand / kern_slope)
  return(emiss)
}

ant_demand <- function(emiss) {
  demand <- (ant_slope * w_base_ant) - (ant_slope * emiss)
  return(demand)
}
ant_demand_inv <- function(demand) {
  emiss <- w_base_ant - (demand / ant_slope)
  return(emiss)
}

vent_demand <- function(emiss) {
  demand <- (vent_slope * w_base_vent) - (vent_slope * emiss)
  return(demand)
}
vent_demand_inv <- function(demand) {
  emiss <- w_base_vent - (demand / vent_slope)
  return(emiss)
}

moj_demand <- function(emiss) {
  demand <- (moj_slope * w_base_moj) - (moj_slope * emiss)
  return(demand)
}
moj_demand_inv <- function(demand) {
  emiss <- w_base_moj - (demand / moj_slope)
  return(emiss)
}

agg_demand <- function(emiss) {
  demand <- -(ant_demand(0) * emiss)/agg_base + ant_demand(0)
  return(demand) # doesnt have steps cuz this is derived from x and why intercept
}

agg_abatement_cost <- function(emiss){
  cost <- (agg_a_slope) * emiss
  return(cost)
}
```

```{r, fig.align="center"}
#plotting demand curves
ggplot() +
  stat_function(fun = kern_demand, aes(color = "Kern")) +
  stat_function(fun = vent_demand, aes(color = "Ventura")) +
  stat_function(fun = moj_demand, aes(color = "Mojave")) +
  stat_function(fun = ant_demand, aes(color = "Antelope")) +
  #stat_function(fun = agg_demand, color = "purple") +
  scale_color_manual(values = c("red","blue","green","orange")) +
  ylim(0,700) +
  xlim(0,300) +
  theme_minimal() +
  labs(title = "Marginal Demand Curves for Water Usage by District", x = "Quantity Water Used (AF)", y = "Marginal Cost of Usage ($/AF)") + theme(legend.title = element_blank(), legend.position = c(.85,.85))
```
The Antelope District is willing to pay the most for the first AF of water as indicated by the red line's y-intercept that is above the other three districts'. 

## Question 3
#### Intervention A
```{r}
reduc_fact = (agg_base-(agg_base-w_abated))/agg_base
reduc_pct = reduc_fact * 100
mult_fact = 1 - reduc_fact

w_base_kern_int_A = w_base_kern * mult_fact
w_base_ant_int_A = w_base_ant * mult_fact
w_base_moj_int_A = w_base_moj * mult_fact
w_base_vent_int_A = w_base_vent * mult_fact
```

```{r}
# cost of intervention to districts and total cost - using demand curves

cost_a_kern <- integrate(kern_demand, w_base_kern_int_A, w_base_kern)$value
cost_a_vent <- integrate(vent_demand, w_base_vent_int_A, w_base_vent)$value
cost_a_ant <- integrate(ant_demand, w_base_ant_int_A, w_base_ant)$value
cost_a_moj <- integrate(moj_demand, w_base_moj_int_A, w_base_moj)$value

total_cost_a = cost_a_kern + cost_a_ant + cost_a_vent + cost_a_moj
```
(1) The stringency of this intervention is a `r round(reduc_pct,1)`% reduction in water usage across all districts, regardless of water usage rates or baseline figures.
(2) The total cost of meeting the 500 AF target through this intervention is $`r format(round(total_cost_a,0),scientific=F)`. 
(3) The cost to each district is as follows:
```{r}
int_a_summary <- c("Cost ($)",round(cost_a_kern,0),round(cost_a_vent,0),round(cost_a_ant,0),round(cost_a_moj,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_a_summary
```
(4) No tax revenue is generated via this intervention. 

#### Intervention B
```{r}
#Aggregate marginal price (P*) of 255 units of abatement is equal to tax 

mpa_agg = w_abated*agg_a_slope #stringency?? P*w dollars/unit abatement 

cost_b_kern = .5*(mpa_agg/kern_slope)*mpa_agg #triangles
cost_b_vent = .5*(mpa_agg/vent_slope)*mpa_agg
cost_b_ant = .5*(mpa_agg/ant_slope)*mpa_agg
cost_b_moj = .5*(mpa_agg/moj_slope)*mpa_agg

total_cost_b = cost_b_kern + cost_b_vent + cost_b_ant + cost_b_moj

tax_rev_b = 500 * mpa_agg 

```
(1) The stringency of this intervention is a $`r round(mpa_agg,0)`/AF tax for water usage across all four districts.
(2) The total cost of meeting the 500 AF target through this intervention is $`r format(round(total_cost_b,0),scientific=F)`.
(3) The cost to each district is as follows:
```{r}
int_b_summary <- c("Cost ($)",round(cost_b_kern,0),round(cost_b_vent,0),round(cost_b_ant,0),round(cost_b_moj,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_b_summary
```
(4) The tax revenue generated via this intervention is the total water usage across all four districts (500 AF) multiplied by the tax from part (1): $`r format(round(tax_rev_b,0),scientific=F)`

#### Intervention C
Regardless of initial allocation, the market will behave according to the equilibrium market principle. The tax from intervention B is equivalent to the price each district would pay for the last unit of abatement (P*). Plugging this value into the inverse of the demand curve functions will yield the quantity of water each district will use after trading is complete. The water used by each district after trade is as follows:

- Kern..........`r round(kern_demand_inv(mpa_agg),0)` AF

- Ventura......`r round(vent_demand_inv(mpa_agg),0)` AF

- Antelope....`r round(ant_demand_inv(mpa_agg),0)` AF

- Mojave.......`r round(moj_demand_inv(mpa_agg),0)` AF

(1) The stringency of this intervention is a `r round(reduc_pct,1)`% reduction in water usage across all districts, regardless of water usage rates or baseline figures. Trading follows this initial allocation of reduction. 
```{r}
# cost of intervention to districts and total cost - using demand curves

cost_c_kern <- integrate(kern_demand, kern_demand_inv(mpa_agg), w_base_kern)$value
cost_c_vent <- integrate(vent_demand, vent_demand_inv(mpa_agg), w_base_vent)$value
cost_c_ant <- integrate(ant_demand, ant_demand_inv(mpa_agg), w_base_ant)$value
cost_c_moj <- integrate(moj_demand, moj_demand_inv(mpa_agg), w_base_moj)$value

total_cost_c = cost_c_kern + cost_c_ant + cost_c_vent + cost_c_moj
```
(2) The total cost of meeting the 500 AF target through this intervention is $`r format(round(total_cost_c,0),scientific=F)`.
(3) The cost to each district is as follows:
```{r}
int_c_summary <- c("Cost ($)",round(cost_c_kern,0),round(cost_c_vent,0),round(cost_c_ant,0),round(cost_c_moj,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_c_summary
```
(4) No tax revenue is generated via this intervention.

## Question 4
Cost of drought to each district under interventions A,B,& C.

**Intervention A:**
```{r}
w_abated = 455

reduc_fact_2 = (agg_base-(agg_base-w_abated))/agg_base
reduc_pct_2 = reduc_fact_2 * 100
mult_fact_2 = 1 - reduc_fact_2

w_base_kern_int_A_2 = w_base_kern * mult_fact_2
w_base_ant_int_A_2 = w_base_ant * mult_fact_2
w_base_moj_int_A_2 = w_base_moj * mult_fact_2
w_base_vent_int_A_2 = w_base_vent * mult_fact_2

#intervention A costs
cost_a_kern_2 <- integrate(kern_demand, w_base_kern_int_A_2, w_base_kern)$value
cost_a_vent_2 <- integrate(vent_demand, w_base_vent_int_A_2, w_base_vent)$value
cost_a_ant_2 <- integrate(ant_demand, w_base_ant_int_A_2, w_base_ant)$value
cost_a_moj_2 <- integrate(moj_demand, w_base_moj_int_A_2, w_base_moj)$value

total_cost_a_2= cost_a_kern_2 + cost_a_ant_2 + cost_a_vent_2 + cost_a_moj_2

int_a_summary_2 <- c("Cost ($)",round(cost_a_kern_2,0),round(cost_a_vent_2,0),round(cost_a_ant_2,0),round(cost_a_moj_2,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_a_summary_2
```

**Intervention B:**
```{r}
mpa_agg = w_abated*agg_a_slope #stringency?? P*w dollars/unit abatement 

cost_b_kern = .5*(mpa_agg/kern_slope)*mpa_agg #triangles
cost_b_vent = .5*(mpa_agg/vent_slope)*mpa_agg
cost_b_ant = .5*(mpa_agg/ant_slope)*mpa_agg
cost_b_moj = .5*(mpa_agg/moj_slope)*mpa_agg

total_cost_b = cost_b_kern + cost_b_vent + cost_b_ant + cost_b_moj

tax_rev_b = 500 * mpa_agg 

int_b_summary_2 <- c("Cost ($)",round(cost_b_kern,0),round(cost_b_vent,0),round(cost_b_ant,0),round(cost_b_moj,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_b_summary_2
```
**Intervention C:**
```{r}
cost_c_kern <- integrate(kern_demand, kern_demand_inv(mpa_agg), w_base_kern)$value
cost_c_vent <- integrate(vent_demand, vent_demand_inv(mpa_agg), w_base_vent)$value
cost_c_ant <- integrate(ant_demand, ant_demand_inv(mpa_agg), w_base_ant)$value
cost_c_moj <- integrate(moj_demand, moj_demand_inv(mpa_agg), w_base_moj)$value

total_cost_c = cost_c_kern + cost_c_ant + cost_c_vent + cost_c_moj

int_c_summary_2 <- c("Cost ($)",round(cost_c_kern,0),round(cost_c_vent,0),round(cost_c_ant,0),round(cost_c_moj,0)) %>% #create vector of values
  t() %>% 
  kable(col.names = c("","Kern","Ventura","Antelope","Mojave")) %>%
  kable_styling(full_width = TRUE)

int_c_summary_2
```
**Preferred Policies:**

- Kern..........Policy A

- Ventura......Policy A

- Antelope....Policy B or C

- Mojave.......Policy B or C

Certain policies may be preferable over others to specific districts because the initial allocation of rights will dictate how many AF a district may have to buy or sell after allocation. If P* is the marginal abatement cost of the final unit of abatement for each district, then each district will have a unique Q* depending on the shape of their abatement curves. Depending on whether the allocation rights of each district fall above or below their unique value of Q* for each policy will dictate whether that policy is preferable to that specific district or not. To generalize for the four regions present, more water-rich regions with flatter abatement curves (Kern & Ventura) prefer a standard percentage reduction because they are more equipped to bear the costs of abatement. Less water-rich regions with steeper abatement curves (Antelope & Mojave) prefer more equitable policies or policies that allow trade because they are not as financially equipped to bear the costs of abating according to a percentage rate.